### 2020-07-16 table_formatting.Rmd

---
title: "Table_formatting"
author: "Yuru"
date: "03/04/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Editting the matlab writen csv table

```{r}
table1 <- read.csv("check1.csv")
colmax <- apply(table1,2,max)
row.names(table1) <- colnames(table1)

GARCH_t12 <- read.csv("GARCH_t12__table.csv")
#row.names(GARCH_t12) <- colnames(GARCH_t12)

GARCH_t3 <- read.csv("GARCH_t3_table.csv")
#row.names(GARCH_t3) <- colnames(GARCH_t3)

ARCH_table <- read.csv("ARCH_table.csv")
ar#row.names(ARCH_table) <- colnames(ARCH_table)

Copula_table <- read.csv("Copula_table.csv")
#row.names(Copula_table) <- colnames(Copula_table)

library(huxtable)
table1_hux <- as_hux(table1,add_colnames = TRUE,add_rownames = TRUE)
table1_hux
table1_hux <- set_number_format(table1_hux,c(2:7),c(2:7),4)

# compare the values in the table to the maximum 
for (columns in 2:7){
for (rows in 2:7) {
  if(identical(table1_hux[rows,columns],colmax[columns]) == 'TRUE'){
    table1_hux <- set_bold(table1_hux,rows,columns,TRUE,byrow = FALSE)
  }
}}

table1_hux

GARCH_t12 <- as_hux(GARCH_t12,add_colnames = TRUE)
GARCH_t3 <- as_hux(GARCH_t3,add_colnames = TRUE)
ARCH_table <- as_hux(ARCH_table,add_colnames = TRUE)
Copula_table <- as_hux(Copula_table,add_colnames = TRUE)
```

```{r}
# set the maximum in each column to bold text
for (columns in 1:5){
  for (rows in 1:5) {
    if (table1_hux[rows+1,columns+1] > table1_hux[rows,columns]){
      max = data.frame(rows+1,columns+1)}
    #else if (table1_hux[rows+1,columns+1] == #table1_hux[rows,columns]){
      #max = data.frame
   # }
    else {max = data.frame(rows,columns)
    }
  table1_hux <- set_bold(table1_hux,max[1,1],max[1,2],TRUE,byrow = FALSE)
  }
  # reset max to be zero dataframe
  max = data.frame(0,0)
}

```

```{r}
# set the maximum in each column to bold text

# find the maximum in each column

max <- matrix(0,nrow=1,ncol=6)
# index <- matrix(nrow=6,ncol = 6)

for (columns in 1:5){
  for (rows in 1:5) {
    if (table1_hux[rows+1,columns+1] > table1_hux[rows,columns]){
      max[1,columns] <- table1_hux[rows+1,columns+1]}

    else {max[1,columns] <- table1_hux[rows,columns]
    }
  }
}

for (columns in 1:5){
  for (rows in 1:5) {
    if (table1_hux[rows+1,columns+1] > table1_hux[rows,columns]){
      max}

    else {max = data.frame(rows,columns)
    }
  table1_hux <- set_bold(table1_hux,max[1,1],max[1,2],TRUE,byrow = FALSE)
  }
  # reset max to be zero dataframe
  max = data.frame(0,0)
}

```

```{r}
# set the maximum in each column to bold text
#（2）
for (columns in 1:5){
#此处循环寻找处整列中的最大值
for (rows in 1:5) {
if (table1_hux[rows+1,columns] > table1_hux[rows,columns]){
#【依据Java的标准】你在循环体内定义的变量max，可能在（1）处用不了，需要在（2）处定义一个变量
max = data.frame(rows+1,columns)}
#else if (table1_hux[rows+1,columns+1] == #table1_hux[rows,columns]){
#max = data.frame
else {max = data.frame(rows,columns)}
#table1_hux <- set_bold(table1_hux,max[1,1],max[1,2],TRUE,byrow = FALSE)
  }
#依据上个循环找出的最大值，再次循环整列来和最大值做比较，相等即为要找的位置
for (rows in 1:6) {
#（1）
temp = data.frame(rows,columns)
if (max == temp){
table1_hux <- set_bold(table1_hux,rows,columns,TRUE,byrow = FALSE)
}
}
max = data.frame(0,0)
}
```

```{r}
# THIS IS THE RIGHT CODE!!!!!
# find the maximum of each column and store this in a 1x6 matrix

colmax <- apply(table1,2,max)

# compare the values in the table to the maximum 
for (columns in 1:6){
for (rows in 1:6) {
  if(table1_hux[rows,columns] >= colmax[columns]){
    table1_hux <- set_bold(table1_hux,rows,columns,TRUE,byrow = FALSE)
  }
}}

table1_hux
```

```{r}
#for (columns in 1:1){
  max1 = -100
  for (rows in 1:5) {
    if(max1 >= table1_hux[rows+1,columns]){
      max1 = max1
    }
    else {max1 = table1_hux[rows,columns]}
  }

for (columns in 1:6) {
  for (rows in 1:6) {
    if (table1_hux[rows,columns] == max1){
    table1_hux <- set_bold(table1_hux,rows,columns,TRUE,byrow = FALSE)
  }
  }
}
table1_hux

  if (table1_hux[rows,columns] == max1){
    table1_hux <- set_bold(table1_hux,rows,columns,TRUE,byrow = FALSE)
  }
#}

```