### 2020-07-31 Draft.Rmd
new intro + update S&P500 series

---
compact-title: no
editor_options:
  chunk_output_type: console
  header-includes: \usepackage{amsmath}
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  message = FALSE,
  warning = FALSE)
```

```{r}
knitr::include_graphics("draft cover page.pdf")
```

\pagebreak

```{r, message=FALSE}
library(formattable)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tsibble)
library(tsibbledata)
library(feasts)
library(normtest)
library(moments)
library(ggthemes)
library(ggpubr)
library(knitr)
library(kableExtra)
```

# Abstract

With correct model specification being an ongoing issue for conventional likelihood-based prediction, scoring rules have been used to produce forecasts that are accurate – in a well-defined sense - in the presence of this misspecification. Based on previous research on this topic, this paper conducts both numerical and empirical analyses to explore the impact of the form and degree of model misspecification on probabilistic forecasts that are optimal under a given score. In particular, the role of optimal forecasts in predicting Value-at-Risk and the VIX Volatility Index, and in optimizing portfolio allocation, is to be investigated.
Results at the current stage have shown that the greater the degree of model misspecification is, the more beneficial is optimal prediction. Empirical analysis shows that after the global financial crisis, predictions that are optimal according to a censored scoring rule that rewards predictive accuracy in a tail are substantially more accurate than the alternatives, including likelihood-based predictions. This indicates the potentially significant contribution to financial analyses of this method.

\pagebreak

# 1. Introduction

This research paper will explore the question: ‘When do optimal probabilistic forecasts work’; with particular attention given to the usefulness of optimal predictions in financial markets. This work builds on recent research on this topic by Loaiza-Maya et al. (2020). Probabilistic forecasts can provide complete information about future uncertainty, which could be more valuable than point and interval forecasts to forecasters. However, both Bayesian and frequentist prediction assume that the underlying model correctly specifies the process that has generated the observations. While in practice, the assumed model underpinning the likelihood function will almost certainly differ from the unknown true data generation process (DGP). This model misspecification problem has been an ongoing issue for conventional likelihood-based prediction.

As an alternative to likelihood-based prediction, scoring rules have been introduced before to produce and evaluate probabilistic forecasts. A variety of proper scoring rules allows users to produce probability predictions based on their interests, and to compare these predictions, as long as they refer to the same forecast quantity (Gneiting & Raftery, 2007). Some recent and previous researches show that forecasts produced based on a given scoring rule can be improved for both Bayesian and frequentist prediction, even in the presence of model misspecification; see Loaiza-Maya, Martin & Frazier (2019), Loaiza-Maya et al. (2020) and (ref 5). Among all these, censored likelihood score (CSL) or focused score (FSR) proposed by (ref 1) captures our attention since its ‘focusing’ feature shows great values in financial risk management. In other words, CSL/FSR allows a forecaster to focus on predicting any region of his/her interests more accurately and therefore, it is expected to make great contributions to risk estimation in finance.

The ‘optimal’ probabilistic forecast has been discussed in Loaiza-Maya et al. (2020). It refers to a predictive probability distribution that is optimal according to a user-specified scoring rule. However, the discussion of whether the optimal forecasts’ performance differs among different types of model misspecification is left to blank, and the financial application of optimal forecasts can be further investigated among different financial instruments.

This research paper extends the work of Loaiza-Maya et al. (2020) to investigate the effects of the form and degree of model misspecification on optimal forecast performance and to conduct empirical analysis on how optimal forecasts perform in Value-at-Risk (VaR), the Volatility Index (VIX) and portfolio optimization applications, using empirical data extending over a period of time that precedes the global financial crisis (GFC) and extends to the latest period in which COVID-19 has impacted on financial markets.

Section 2 documents how scoring rules work in producing and evaluating the optimal forecasts under both simulation and empirical settings. The background and notation of scoring rules used in this paper are provided at the start. Besides, the simulation analysis investigates the effects of different forms and degrees of model misspecification on optimal forecast performance using data simulated from the (generalized) autoregressive conditional heteroscedasticity ((G)ARCH) and copula models. The assumed predictive model is fixed as ARCH (1) for the purpose of manipulating the degree of model misspecification. The performance of optimal forecasts in practice is also shown in this section, conducted with S&P500 daily log return data. In section 3, we look at the financial application of scoring rules in VaR using both simulated and S&P500 observations with same methodology adopted in section 2. The analysis of VIX is provided in section 4, and the predictive model is HAR-RV model with different error term specifications. We conclude and discuss any possible caveats in section 5.

# 2. Scoring rule in prediction
## 2.1 Overview and notation

Scoring rules can be used in both estimation and prediction. In estimation, an optimum estimator $\hat \theta$ can be obtained by maximizing a scoring rule $S_n$. That is, $\hat \theta_n=arg \max_\theta S_n(\theta)$. $\hat \theta \to \theta$ as $n \to \infty$ under asymptotic situations. While in prediction, scoring rules assign a numerical value to density forecasts, addressing both sharpness and calibration. For a positively oriented score, higher value will be assigned to a better forecast between two competing candidates. A variety of proper scoring rules allows users to produce probability predictions based on their interests, and to compare these predictions, as long as they refer to the same forecast quantity. Forecasts candidates are often compared by their average scores. Therefore, the crucial importance of the propreity of scoring rules must be emphasized.Gneiting & Raftery, (2007) illustrate the importance of propreity via a case study in weather forecasts and point out that improper scoring rules will assign higher average score to an incorrect density forecast. A scoring rule is said to be proper if $S(Q,Q) \geq S(P,Q)$ for all P and Q, and it is strictly proper if $S(Q,Q) = S(P,Q)$ only happens when P = Q. P and Q are predictive distributions and Q is the best forecast given all the available information. S(P,Q) means the expected value of S(P,$\cdot$) under Q (Gneiting & Raftery, 2007).

Some commonly used proper scoring rules are the logarithmic score (LS), the continuously ranked probability score (CRPS) and censored likelihood score (CSL) or focused score (FSR) in a sense that CSL is used for focusing on a certain region(s) of densities.

The logarithmic score is defined as (1) where $p_t$ is the density forecasts. It is a local strictly proper scoring rule, which means it will assign a higher score to the correct probabilistic forecast, and it is superior to quadratic and spherical scoring rules when the rank ordering is important or the impact of nonlinear utility function used is a concern for forecasters (Bickel, 2007). 
\begin{align}
S_{LS}(p_t, y_{t+1}) = log[p_t(y_{t+1})]
\end{align}
However, logarithmic score is criticized by its unboundness and its local property. Its useness is also questioned in (Bernardo, 1979), since he shows that every proper scoring rule is equivalent to logarithmic score. Gneiting & Raftery, (2007) argue that logarithmic score is insensitive to distance and will not reward predictions that are close to but not identical to the materialized one. Therefore, Gneiting & Raftery, (2007) propose the continuously ranked probability score, which is sensitive to distance and is defined as in (2) where P is the cumulative distriubtion function, I is the indication function, and x is the materialized events. 
\begin{align}
CRPS(P,x)=-\int_{-\infty}^{\infty}(P(y)-I(y\geq x))^2dy
\end{align}
The formula can be simplifed to (3) if the prediction distribution is Gaussian with mean = $\mu$ and variance = $\sigma^2$.
\begin{align}
CRPS(N,x)=\sigma[\frac{1}{\sqrt{\pi}}-2\phi(\frac{x-\mu}{\sigma})-\frac{x-\mu}{\sigma}(2\Phi(\frac{x-\mu}{\sigma})-1)]
\end{align}
where $\phi$ and $\Phi$ are the probability density function and cumulative density function of the Gaussian predictive distribution.

Continuously ranked probability score is defined as a negatively oriented scoring rule, but it can be easily transformed to positively oriented score for convenient comparison among scoring rules as shown in (4). 
\begin{align}
CRPS^*(N,x)=\sigma[-\frac{1}{\sqrt{\pi}}+2\phi(\frac{x-\mu}{\sigma})+\frac{x-\mu}{\sigma}(2\Phi(\frac{x-\mu}{\sigma})-1)]
\end{align}
Both logarithmic score and continuously ranked probability score are used for producing and evaluating the entire predictive density. In terms of accurately predicting a certain region of a distribution, they are typically used together with weighted likelihood (Gneiting & Ranjan, 2011). Diks, Panchenko & van Dijk, (2011) propose a new proper scoring rule, censored likelihood score, which allows users to assess the forecasts only in the interested region(s) instead of using weights to emphasize a particular part of the entire density forecast. It is defined as in (5) where $p_t$ is the predictive probability density function, $A_t$ is the interested region(s) and $A_t^c$ is the complement of $A_t$.
\begin{align}
S_{CSL}(p_t, y_{t+1})=I(y_{t+1} \in A_t)log[ p_t(y_{t+1})]+I(y_{t+1} \in A_t^c)log[\int_{A_t^c} p_t(s)ds]
\end{align}
[ref 5] prove that weighted density forecasts based on optimizing the censored likelihood score outperform those on logarithmic score and continuously ranked probability score.

## 2.2 Illustration example 
### 2.2.1 Simulation overview

Density forecasts provide more complete information of uncertainty of a prediction than point and interval forecasts. Moreover, the tails of a density have a great importance in finance. It forms the foundation of risk management, such as Value-at-Risk (VaR), and asset allocation. It can also be used in derivatives pricing. (reference ?) --- find some papers.

[ref 5] provide empirical results showing that using censored likelihood score can improve 99% Value-at-Risk estimation. Moreover, Loaiza-Maya, Martin & Frazier, (2019) propose a new method, Focused Bayes, in Bayeisian prediction by replacing the conventional likelihood function with the censored likelihood score. Their results show that Focused Bayes outperform exact Bayes which is using logarithmic score for updating the prior probability. Their work also points out the degree of model misspecification play a role in the performance of Focused Bayes.

Loaiza-Maya et al., (2020) extend the above discussion to frequentist probability forecasting. They address questions about when we can/cannot benefit from optimal probabilistic forecasts out of sample. A predictive probability distribution is said to be optimal if it is optimal based on a user-specified scoring rule, not generally speaking.

Building on the previous research in this field, we conduct a numerical analysis first in section 2.2.2 and 2.2.3 to further investigate the effects of the form and degree of model misspecification on optimal forecast performance using data simulated from the (generalized) autoregressive conditional heteroscedasticity ((G)ARCH) and copula models.

### 2.2.2 Simulation design

In order to prepare the financial application analysis in following sections of this paper, we simulate a time series dependnet variable $y_t$ that mimic the behavior of financial returns and volatility. Specifically, GARCH models can capture the volatility clustering and serial dependence, and the negative marginal skewness will be introducted using an inversion copula with degree of skewness controlled by the shape parameter showed below. In order to manipulate the degree and form of model misspecification, we fix the underlying model to a simple and clearly misspecified ARCH(1) model.

Specific simulation scenarios are listed in following table ($t_\nu$ indicates a Student-t distribution with $\nu$ degrees of freedom).

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Simulation design}} \\ \hline
          & \textbf{Scenario (i)} & \textbf{Scenario (ii)} \\ \hline
\textbf{True DGP} & $y_t = \sigma_t \epsilon_t$ & copula \\
& $\sigma_t^2 = 1 + 0.2y_{t-1}^2+0.7\sigma_{t-1}^2$ & Shape parameter = {0, -3, -5}\\ & $\epsilon_t \sim (\frac{\nu-2}{\nu})^0.5 * t_\nu$ & \\ & $\nu \in (3, 12, 10000)$ & \\ \hline

\textbf{Assumed Model} & $y_t = \mu +\sigma_t \epsilon_t$ & $y_t = \mu +\sigma_t \epsilon_t$\\ & $\sigma_t^2 = \alpha_0 + \alpha_1 \sigma_{t-1}^2$ & $\sigma_t^2 = \alpha_0 + \alpha_1 \sigma_{t-1}^2$\\ & $\epsilon_t \sim N(0,1)$ & $\epsilon_t \sim N(0,1)$\\ \hline
\end{tabular}
\end{table}

The tails of density are important for risk management in finance so that we decide to apply focused scoring rule (FSR) to 10%, 20%, 80% and 90% tails of the distribution, corresponding to short and long portfolios in finance respectively. The methodology in Loaiza-Maya et al., (2020) and Loaiza-Maya, Martin & Frazier, (2019) is adopted. We conduct each simulation scenario as follows (Loaiza-Maya et al., 2020):

  1.	Generate T = 6000 observations for $y_t$ from the true DGP

  2.	Use $y_{1:1000}$ to estimate $\boldsymbol{\hat \theta}$ in the assumed predictive model based on the positively-oriented score $S_i$ where $\boldsymbol{\hat \theta} = \{\hat \mu, \hat \alpha_0, \hat \alpha_1\}$

\begin{align}
{\hat \theta^{[i]}}:=arg\max_{\theta \in \Theta}\bar S_i({\theta})
\end{align}

\begin{align}
\bar S_i(\theta):=\frac{1}{T-(\tau+1)}\sum^{T-\tau}_{t=2}S(P_{\theta}^{t-1},y_t)
\end{align}

  3.	Produce the one-step-ahead predictive $P^{t-1}_{\hat \theta_1^{[i]}}$ And compute the out-of-sample score using $S_j$, where $S_i$ and $S_j$ refer to (8)(9)(10):

\begin{align}
S_{LS}(P_\theta^{t-1},y_t)=lnp(y_t|\mathcal{F}_{t-1},\theta)
\end{align}

\begin{align}
S_{CRPS}^*(P_\theta^{t-1},y_t)=\int^{+\infty}_{-\infty}[P(y|\mathcal{F}_{t-1},\theta)-I(y \geq y_t)]^2dy
\end{align}

\begin{align}
S_{FSR}(P_\theta^{t-1},y_t)=lnp(y_t|\mathcal{F}_{t-1},\theta)I(y_t \in A)+[ln\int_{A^c}{p(y|\mathcal{F}_{t-1},\theta)dy}]I(y_t \in A^c)
\end{align}

4. Expand estimation window by one observation and repeat step 2-3 with $\tau = T-1000$ times and compute the average scores: 
\begin{align}
\bar S_j(\hat \theta^{[i]})=\frac{1}{\tau}\sum^T_{t=T-\tau+1}S_j(P_{\hat \theta^{[i]}}^{t-1},y_t)
\end{align}

### 2.2.3 Simulation results
#### 2.2.3.1 Average score tables

Loaiza-Maya et al., (2020) introduce the concepts of ‘coherence’ and ‘strict coherence’ for conveniently documenting the optimal forecasts’ performance. Coherence means when the optimal probabilistic forecast based on a given score is superior, or at least performs the same as, alternative forecasts according to the same score. Strict coherence is when the optimal prediction is strictly preferable given that score. Let $\boldsymbol{\hat \theta}_1$ and $\boldsymbol{\hat \theta}_2$ be the optimizers based on $S_1, S_2$, then the prediction is said to coherent if 
\begin{align}
\frac{1}{\tau}\sum_{t=T-\tau+1}^{T}S_1(P^{t-1}_{\boldsymbol{\hat \theta_1}},y_t) \geq \frac{1}{\tau}\sum_{t=T-\tau+1}^{T}S_1(P^{t-1}_{\boldsymbol{\hat \theta_2}},y_t)
\end{align}

\begin{align}
\frac{1}{\tau}\sum_{t=T-\tau+1}^{T}S_2(P^{t-1}_{\boldsymbol{\hat \theta_1}},y_t) \leq \frac{1}{\tau}\sum_{t=T-\tau+1}^{T}S_2(P^{t-1}_{\boldsymbol{\hat \theta_1}},y_t)
\end{align}

It is said to be strictly coherent if the equality doesn't hold in (11)(12).

\textbf{[need to mention some results from opt paper - one sentence]}

The first column of following tables shows $S_i$ that we used to produce predictions and the first row shows $S_j$, which is used for forecasts evaluation. The bolded value(s) is the maximum(s) in each column. We use all positively oriented scoring rules, and therefore, the column maximum(s) indicate(s) the optimal prediction(s) based on the score indicated by the column name.

```{r}
# Editting the matlab written csv files

GARCH_t12 <- read.csv("GARCH_t12__table.csv")
GARCH_t12 <- rename(GARCH_t12, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
GARCH_t12$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

GARCH_t3 <- read.csv("GARCH_t3_table.csv")
GARCH_t3 <- rename(GARCH_t3, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
GARCH_t3$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

ARCH_table <- read.csv("ARCH_table.csv")
ARCH_table <- rename(ARCH_table, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
ARCH_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

Copula5_table <- read.csv("Copula5_table.csv")
Copula5_table <- rename(Copula5_table, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
Copula5_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

Copula3_table <- read.csv("Copula3_table.csv")
Copula3_table <- rename(Copula3_table, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
Copula3_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

Copula0_table <- read.csv("Copula0_table.csv")
Copula0_table <- rename(Copula0_table, "In-sample optimizers" = Row, "FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
Copula0_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

```

```{r}
ARCH_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 1: the true DGP is ARCH(1)"=7))
```

```{r}
GARCH_t12 %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 2: the true DGP is GARCH(1,1) with degree of freedom = 12"=7))
```

```{r}
GARCH_t3 %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 3: the true DGP is GARCH(1,1) with degree of freedom = 3"=7))
```

The results from Table 1-3 show one particular type of model misspecification. That is, the underlying model ARCH(1) cannot capture the volatility clustering feature of the true data generation process (DGP), thus misspecification. Besides, with the degree of freedom increases, the data generated from GARCH(1,1) model is more like data that are from ARCH(1) and therefore, the degree of model misspecification decreases. When degree of freedom is large enough, we consider ARCH(1) as a correct specified model as shown in Table 1.

From Table 3. we could observe strict coherence when degree of model misspecification is high and coherence in Table 1 which is stated as correct specification case. These results are consistent with conclusions from (Loiaza-Maya et al., 2020). It simply means that we can improve prediction accuracy in the presence of model misspecification ?? , and gain more from using optimal forecasts method while the degree of model misspecification is high. In other words, we are expected to gain more from using focused score when we observe 'strict coherence' while comparing two forecasts.

```{r}

Copula0_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 4: the true DGP is copula with shape parameter = 0"=7))
```

```{r}
Copula3_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 5: the true DGP is copula with shape parameter = -3"=7))
```

```{r}
Copula5_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 6: the true DGP is copula with shape parameter = -5"=7))
```

\textbf{[need copula paper citation here to explain more]}

Table 4-6 show another type of model misspecification where the underlying model cannot capture the asymmetric feature of the true DGP. The results give similar conclusions to Table 1-3. It proves that optimal forecasts perform in the same way even when we have different types of model mispecification.

#### 2.2.3.2 Score traces plots

As stated in previous literature, focused score tends to outperform CRPS and logarithmic score in predicting tails. Therefore, we provide score trace plots to exhibit the comparison of LS, CRPS and FSR’s performance in predicting the 10 percentile and 90 percentile in figure 1 and 2. In the left panel, it shows the performance of each score evaluated by FSR 10%. That is, if the average score value of score A used for estimation is higher than score B, then score A predicts the 10 percentile of the distribution more accurately than score B. The right panel shows the same thing but is based on performance of predicting 90 percentile. From figure 1 and 2, we could see that FSR makes no difference if the degree of model misspecification is very small but as it increases, FSR starts to outperform CRPS and MLE (LS) in predicting tails. The more misspecified the model is, the more benefits can we gain from using FSR. It is beneficial for econometricians to improve forecast accuracy without working so hard on finding a ‘correct’ model in the fact that in practice, a correct model does not exist.

```{r}
# need to adjust the size. 
knitr::include_graphics("figure 1.pdf")
```

```{r}
knitr::include_graphics("figure 2.pdf")
```

Moreover, the predictive density plots (Appendix 1) in figure 3 to 8, further support our conclusions drawn from the average score tables and score trace plots. They show that predictions obtained from optimizing focused scores aim to match the shape of tails of the true distribution and therefore, outperform predictions from using LS.

## 2.3 Empirical analysis

From the results in section 2.2, we can conclude that optimal forecasts based on focused score can improve the prediction accuracy in tails and its effect is clearest when the degree of model misspecification is high and does not vary much when we have different types of model misspecification. In section 2.3, we apply this method to an empirical setting and investigate how it performs in practice.

The data used in the analysis is the continuously compounded daily log returns of S&P500 listed in U.S.A financial market over a time period that precedes the 2008 Global financial crisis (GFC) and extends to the latest period in which COVID-19 has impacts on financial market. The methodology used here is the same as in the section 2.2.2 with initial training sample size set to be 1000 and 5000 out-of-sample observations.

### 2.3.1 Empirical analysis results

```{r}
# data cleaning and import 
sp <- read.csv("^GSPC.csv")
sp <- sp %>% dplyr::select("Date","Adj.Close")

sp$log.returns <- c(NA,diff(log(sp$Adj.Close)))
logret <- sp[-c(1:188),c(1,3)] # delete the adjusted price col, only keep 6000 obs

# export the dataset to csv. including only trading days and log returns 
temp <- logret[,c(2)]
temp <- data.frame(temp)
write_csv(temp,"sp500log.csv")
```

```{r, results='hide'}
# descriptive statistics 
summary <- summary(logret$log.returns) # ----> need to be tabulated 
# range(logret$log.returns)
# some test results aginst normal dist h_null
# JB test
jb.norm.test(logret$log.returns)
# Ljung box test
Box.test(logret$log.returns,lag = 2,type = "Lj",fitdf = 1)

sk <- skewness(logret$log.returns)
kt <- kurtosis(logret$log.returns)
```

```{r}
sum <- data.frame(Stock="S&P500",Min=as.numeric(summary[1]),Median=as.numeric(summary[3]),Mean=as.numeric(summary[4]),Max=as.numeric(summary[5]),Skewness=sk,Kurtosis=kt,JB.Test=28270,LB.Test=61.489)
sum %>% 
  kable("latex",booktabs = T) %>% 
  kable_styling("bordered") %>% 
  add_header_above(c("Table ?: Descriptive statistics"=9))
```

```{r,message=FALSE,fig.width=10,fig.height=5}

# time series plot
logret$Date <- as.Date(logret$Date,format = "%Y-%m-%d")

p1 <- logret %>% ggplot(aes(x = Date, y = log.returns)) +
  geom_line()+theme_classic()+labs(x="Year",y="Daily log returns")+
  #ggtitle("Figure 9: S&P500 daily log returns")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  scale_y_continuous(labels = scales::percent) + 
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=-0.12,label="Global Financial Crisis",color="red",size=3)+
  annotate(geom = 'text',x=as.Date("2020-04-01"),y=-0.12,label="COVID-19",color="red",size=2)+
  theme_clean()
  #+labs(caption = "Figure 9: the time series plot of S&P500 daily log returns")

# histogram plots
p2 <- logret %>% ggplot()+geom_histogram(aes(x=log.returns))+xlim(-0.13,0.13)+
  #ggtitle("Figure 10")+
  scale_x_continuous(labels = scales::percent)+theme_clean()
  #+labs(caption = "Figure 10: the distribution of S&P500 log returns is asymmetric")

# ACF plots to see the autocorrelation 
logret <- logret %>% 
  mutate(trading_day = row_number()) %>%
  as_tsibble(index = trading_day,regular = TRUE)

p3 <- logret %>% ACF((log.returns)^2) %>% autoplot() + 
  #ggtitle("Figure 11:Squared log returns autocorrelation")+
  xlab("number of lags")+ylab("ACF for variance")+theme_clean()
# +labs(caption = "Figure 11: the autocorrelation plot shows there is a strong serial correlation in volatility of S&P500")

figure <- ggarrange(p1,                                                 
          ggarrange(p2, p3, ncol = 2, labels = c("B", "C")), 
          nrow = 2, 
          labels = "A"                                        
          ) 
annotate_figure(figure, 
                top = text_grob("Descriptive figures",face = "bold"),
                fig.lab = "Figure 11",fig.lab.face = "bold"
                )
```

The descriptive statistics shows that the distribution of S&P500 log returns are negatively skewed and have fat tails and strong serial correlation. Besides, during the GFC and COVID-19 pandemic periods, the volatility is very high due to high uncertainty.

The reality is always the most complicated and the underlying ARCH(1) should have a high degree of model misspecification. Therefore, we should expect that predictions produced by optimizing FSR do a better job in predicting two tails of the distribution. The results of Table 9 support our conjecture. However, the difference between using FSR10(80) and FSR20(90) is very small.

```{r}
sp_table <- read.csv("sp_table.csv")
sp_table <- rename(sp_table, "In-sample optimizers" = Row,"FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
sp_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

sp_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table 9: Average scores for S&P500"=7))
```

Unlike simulation, the score trace plots show how different scoring rules perform in real time points which correspond to some important events. In figure 12, one interesting phenomenon is that after the GFC, focused score performs much better than CRPS and MLE in terms of predicting both upper and lower tails. And this 'jump' effect continues until the very recent date. It indicates that optimal probabilistic forecasts might predict financial markets better in and after times of stock market turbulance, which could represent a great contribution to the existing financial analyses of risks and returns.

```{r,fig.width=10,fig.height=5}
# to compress all the calculated plotting average scores together in the same df 
trace_fsr10 <- read.csv("fsr10.csv")
trace_fsr20 <- read.csv("fsr20.csv")
trace_10ls <- read.csv("fsr10_ls.csv")
trace_10crps <- read.csv("fsr10_crps.csv")

fsr10 <- data.frame(trace_fsr10,trace_fsr20,trace_10ls,trace_10crps)

# add back dates to the fsr10 df, initial step = 1000

fsr10 <- fsr10 %>% mutate(Date = logret$Date[1001:6000])

# plot scores --- lower tail 
fsr10$Date <- as.Date(fsr10$Date,format = "%Y-%m-%d")

color1 <- c("FSR10"="red","FSR20"="black","MLE"="purple","CRPS"="green")

fsr10plot <- fsr10 %>% ggplot(aes(x=Date)) +
  geom_line(aes(y = FSR10,color = "FSR10"),size=0.25)+
  geom_line(aes(y = FSR20,color = "FSR20"),size=0.25)+
  geom_line(aes(y = MLE,color = "MLE"),size=0.25)+
  geom_line(aes(y = CRPS,color = "CRPS"),size=0.25)+
  labs(x="Year",y="Average Scores",color="legend")+
  #ggtitle("Figure 12: S&P500 Score traces-lower tail")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  scale_color_manual(values = color1)+
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=0.08,label="Global Financial Crisis",color="darkred")+
  theme_clean()

# plot scores --- upper tail 
trace_fsr90 <- read.csv("fsr90.csv")
trace_fsr80 <- read.csv("fsr80.csv")
trace_90ls <- read.csv("fsr90_ls.csv")
trace_90crps <- read.csv("fsr90_crps.csv")

fsr90 <- data.frame(trace_fsr90,trace_fsr80,trace_90ls,trace_90crps)

# add back dates to the fsr10 df, initial step = 1000

fsr90 <- fsr90 %>% mutate(Date = logret$Date[1001:6000])

# plot scores --- lower tail 
fsr90$Date <- as.Date(fsr90$Date,format = "%Y-%m-%d")

color2 <- c("FSR90"="red","FSR80"="black","MLE"="purple","CRPS"="green")

fsr90plot <- fsr90 %>% ggplot(aes(x=Date)) +
  geom_line(aes(y = FSR90,color = "FSR90"),size=0.25)+
  geom_line(aes(y = FSR80,color = "FSR80"),size=0.25)+
  geom_line(aes(y = MLE,color = "MLE"),size=0.25)+
  geom_line(aes(y = CRPS,color = "CRPS"),size=0.25)+
  labs(x="Year",y="",color="legend")+
  #ggtitle("Figure 12: S&P500 Score traces-upper tail")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  scale_color_manual(values = color2)+
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=-0.05,label="Global Financial Crisis",color="darkred")+
  theme_clean()

# put two plots side by side 
ggarrange(fsr10plot,fsr90plot,ncol=2,legend = "bottom",labels = c("Figure 12(a) Lower Tail","Figure 12(b) Upper Tail"))

```

# 3. Finanical application: Value-at-Risk
## 3.1 Overview of Value-at-Risk

Value-at-Risk (VaR) is a commonly used risk measurement in finance. It can be obtained from a probability density function of returns over a chosen investment horizon. It is the p-quantile of a predictive distribution of a portfolio’s profit/loss over a holding period, according to a given confidence interval $\alpha$ (Dowd, 2005).
\begin{align}
VaR = -q_p \\p=1-\alpha
\end{align}
It indicates that investors will not lose more than the amount as shown in VaR. It represents the worst level of the investments with $\alpha$ confidence level. There are different methods to estimate VaR, but it is clear that predictive density plays an important role here. Therefore, a more accurate prediction can contribute to improving the accuracy of VaR estimation.

[ref 5] conduct combining density forecasts using focused scoring rules, and the results show weighted density forecasts based on optimizing the focused score outperform those based on optimizing CRPS or LS and improve the accuracy of 99% Value-at-Risk estimates.

We implement VaR analysis by using VaR backtesting method. It evaluates the prediction accuracy with the proportion of exceedances ($y_t < VaR_t$). That is, an accurate estimation of the p% VaR is observed if the proportion of exceedances equals to p%. We illustrate how optimal forecasts perform in estimating VaR by using simulation data from section 2 first and expect outperformance of focused scoring rules.

## 3.2 Value-at-Risk for simulation analysis

The following tables include VaR estimations at \{10%, 20%, 80%, 90%\} corresponding to the 10% and 20% loss of long and short portfolios.
```{r}
copula5var <- read.csv("copula5var.csv")

VaR_0.1 <- as.numeric(copula5var[,c(1,5,9,13,17,21)])
VaR_0.2 <- as.numeric(copula5var[,c(2,6,10,14,18,22)])
VaR_0.8 <- as.numeric(copula5var[,c(3,7,11,15,19,23)])
VaR_0.9 <- as.numeric(copula5var[,c(4,8,12,16,20,24)])

Optimizers <- c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

df <- data.frame(Optimizers,VaR_0.1,VaR_0.2,VaR_0.8,VaR_0.9)
df <- rename(df,"VaR at 10\\%"=VaR_0.1, "VaR at 20\\%"=VaR_0.2, "VaR at 80\\%"=VaR_0.8, "VaR at 90\\%"=VaR_0.9)

df %>% 
  mutate(
    `VaR at 10\\%` = cell_spec(format(`VaR at 10\\%`,4),"latex",bold = ifelse(abs(`VaR at 10\\%`-0.1) == min(abs(`VaR at 10\\%`-0.1)),TRUE,FALSE)),
    `VaR at 20\\%` = cell_spec(format(`VaR at 20\\%`,4),"latex",bold = ifelse(abs(`VaR at 20\\%`-0.2) == min(abs(`VaR at 20\\%`-0.2)),TRUE,FALSE)),
    `VaR at 80\\%` = cell_spec(format(`VaR at 80\\%`,4),"latex",bold = ifelse(abs(`VaR at 80\\%`-0.8) == min(abs(`VaR at 80\\%`-0.8)),TRUE,FALSE)),
    `VaR at 90\\%` = cell_spec(format(`VaR at 90\\%`,4),"latex",bold = ifelse(abs(`VaR at 90\\%`-0.9) == min(abs(`VaR at 90\\%`-0.9)),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F,booktabs=T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c(" ","Out-of-sample exceedances"=4)) %>%
  add_header_above(c("Table 7: VaR when true DGP is copula (shape = -5)"=5))
```

```{r}
garch3var <- read.csv("garch3var.csv")

VaR0.1 <- as.numeric(garch3var[,c(1,5,9,13,17,21)])
VaR0.2 <- as.numeric(garch3var[,c(2,6,10,14,18,22)])
VaR0.8 <- as.numeric(garch3var[,c(3,7,11,15,19,23)])
VaR0.9 <- as.numeric(garch3var[,c(4,8,12,16,20,24)])

df <- data.frame(Optimizers,VaR0.1,VaR0.2,VaR0.8,VaR0.9)
df <- rename(df,"VaR at 10\\%"=VaR0.1, "VaR at 20\\%"=VaR0.2, "VaR at 80\\%"=VaR0.8, "VaR at 90\\%"=VaR0.9)

df %>% 
  mutate(
    `VaR at 10\\%` = cell_spec(format(`VaR at 10\\%`,4),"latex",bold = ifelse(abs(`VaR at 10\\%`-0.1) == min(abs(`VaR at 10\\%`-0.1)),TRUE,FALSE)),
    `VaR at 20\\%` = cell_spec(format(`VaR at 20\\%`,4),"latex",bold = ifelse(abs(`VaR at 20\\%`-0.2) == min(abs(`VaR at 20\\%`-0.2)),TRUE,FALSE)),
    `VaR at 80\\%` = cell_spec(format(`VaR at 80\\%`,4),"latex",bold = ifelse(abs(`VaR at 80\\%`-0.8) == min(abs(`VaR at 80\\%`-0.8)),TRUE,FALSE)),
    `VaR at 90\\%` = cell_spec(format(`VaR at 90\\%`,4),"latex",bold = ifelse(abs(`VaR at 90\\%`-0.9) == min(abs(`VaR at 90\\%`-0.9)),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs=T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c(" ","Out-of-sample exceedances"=4)) %>%
  add_header_above(c("Table 8: VaR when true DGP is GARCH with degree of freedom = 3"=5))
```

The values showed in these tables refer to the out-of-sample proportion of exceedances over the nominal VaR level indicated by the column name. The closer the proportion of exceedances to nominal VaR levels, the more accurate are VaR estimations, and thus the more accurate are the optimal probablistic forecasts. We can see that in two different forms of model misspecification displayed in Table 7 and 8, focused score can provide the most accurate VaR estimation in the upper tail, while in lower tail, it is not short of advantages by too much compared with CRPS. These results help us understand what to expect in empirical VaR analysis.

## 3.3 Value-at-Risk for empirical analysis

We now implement the VaR analysis to continuously compounded daily log returns of S&P500 in U.S.A financial market. Learning from score trace plots in section 2.3.1, the 2008 GFC is the turning point of the optimal forecasts performance, especially for focused scores. Therefore, we separate the VaR analysis to two different time periods, before GFC and after GFC.

```{r}
spvar <- read.csv("spvar1_table.csv")

VaR_0.1 <- as.numeric(spvar[,c(1,5,9,13,17,21)])
VaR_0.2 <- as.numeric(spvar[,c(2,6,10,14,18,22)])
VaR_0.8 <- as.numeric(spvar[,c(3,7,11,15,19,23)])
VaR_0.9 <- as.numeric(spvar[,c(4,8,12,16,20,24)])

df <- data.frame(Optimizers,VaR_0.1,VaR_0.2,VaR_0.8,VaR_0.9)
df <- rename(df,"VaR at 10\\%"=VaR_0.1, "VaR at 20\\%"=VaR_0.2, "VaR at 80\\%"=VaR_0.8, "VaR at 90\\%"=VaR_0.9)

df %>% 
  mutate(
    `VaR at 10\\%` = cell_spec(format(`VaR at 10\\%`,4),"latex",bold = ifelse(abs(`VaR at 10\\%`-0.1) == min(abs(`VaR at 10\\%`-0.1)),TRUE,FALSE)),
    `VaR at 20\\%` = cell_spec(format(`VaR at 20\\%`,4),"latex",bold = ifelse(abs(`VaR at 20\\%`-0.2) == min(abs(`VaR at 20\\%`-0.2)),TRUE,FALSE)),
    `VaR at 80\\%` = cell_spec(format(`VaR at 80\\%`,4),"latex",bold = ifelse(abs(`VaR at 80\\%`-0.8) == min(abs(`VaR at 80\\%`-0.8)),TRUE,FALSE)),
    `VaR at 90\\%` = cell_spec(format(`VaR at 90\\%`,4),"latex",bold = ifelse(abs(`VaR at 90\\%`-0.9) == min(abs(`VaR at 90\\%`-0.9)),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs=T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c(" ","Out-of-sample exceedances"=4)) %>%
  add_header_above(c("Table 11: VaR for S&P500 before GFC"=5))

# formattable(df,list(VaR_0.1 = max_bold1,VaR_0.2 = max_bold2,VaR_0.8 = max_bold8,VaR_0.9 = max_bold9))
```

```{r}
spvar <- read.csv("spvar2_table.csv")
spvar <- round(spvar,digits = 4)

VaR_0.1 <- as.numeric(spvar[,c(1,5,9,13,17,21)])
VaR_0.2 <- as.numeric(spvar[,c(2,6,10,14,18,22)])
VaR_0.8 <- as.numeric(spvar[,c(3,7,11,15,19,23)])
VaR_0.9 <- as.numeric(spvar[,c(4,8,12,16,20,24)])

df <- data.frame(Optimizers,VaR_0.1,VaR_0.2,VaR_0.8,VaR_0.9)
df <- rename(df,"VaR at 10\\%"=VaR_0.1, "VaR at 20\\%"=VaR_0.2, "VaR at 80\\%"=VaR_0.8, "VaR at 90\\%"=VaR_0.9)

df %>% 
  mutate(
    `VaR at 10\\%` = cell_spec(format(`VaR at 10\\%`,4),"latex",bold = ifelse(abs(`VaR at 10\\%`-0.1) == min(abs(`VaR at 10\\%`-0.1)),TRUE,FALSE)),
    `VaR at 20\\%` = cell_spec(format(`VaR at 20\\%`,4),"latex",bold = ifelse(abs(`VaR at 20\\%`-0.2) == min(abs(`VaR at 20\\%`-0.2)),TRUE,FALSE)),
    `VaR at 80\\%` = cell_spec(format(`VaR at 80\\%`,4),"latex",bold = ifelse(abs(`VaR at 80\\%`-0.8) == min(abs(`VaR at 80\\%`-0.8)),TRUE,FALSE)),
    `VaR at 90\\%` = cell_spec(format(`VaR at 90\\%`,4),"latex",bold = ifelse(abs(`VaR at 90\\%`-0.9) == min(abs(`VaR at 90\\%`-0.9)),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs=T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c(" ","Out-of-sample exceedances"=4)) %>%
  add_header_above(c("Table 12: VaR for S&P500 after GFC"=5))

# formattable(df,list(VaR_0.1 = max_bold1,VaR_0.2 = max_bold2,VaR_0.8 = max_bold8,VaR_0.9 = max_bold9))
```

The results from Table 11 and 12 show that the focused score perform well enough in terms of estimating VaR for both long and short portfolios of S&P500, but also CRPS is a very robusted scoring rule which can provide relatively accurate predictions. Moreover, in Table 12, focused score seems to perform better in 20% and 80% which have higher probability mass. If we think of the marginal distribution of S&P500 estimated by histogram in section 2.3.1, the above results could be a consequence of the very sparsed marginal distribution, which increases the difficulty of 'focusing' on the 10% and 90% tails.

# 4. Empirical analysis in VIX
## 4.1 Background and notation

Time-varying volatility is an important part in finance modelling and an accurate prediction will give investors a better understanding of the risks they take on. GARCH-type models used in section 2 and 3 are conditional deterministic. That is, $\sigma_t^2$ is a deterministic function of given past returns. They neglect the fact that volatility in stock markets and derivative markets can have long memory, which means it slowly reverts to its long-run mean. They do not exploit the information in intraday data. Therefore, it is better off if we model volatility in a continuous-time model where $lnP_t$ is allowed to vary continuously over time.

The realized volatility (RV) approach to modelling volatility allows us to exploit the information in intraday data, consider long memory feature, sudden jumps in the market and market microstructure. When the price follows the process as:
\begin{align}
d\ln(P_t)=\mu_tdt+\sigma_tdw_t
\end{align}

RV is a direct estimate of integrated volatility (IV) for $lnP_t$ based on continuously recorded observations of $P_t$ over the day.
\begin{align}
RV_t = \sum^{1/\Delta t}_{j=1}r^2_{t-1+j\Delta t} \xrightarrow{p} \int_{t-1}^t\sigma_s^2ds = IV_t \ as\ \Delta t \to 0
\end{align}
When considering the sudden jumps, that is, the price follows the process in (18). RV is also a consistent estimate of quadratic variation (QV) where $\kappa_s$ represents the sudden jump on day t and $q_t$ is the jump occasions.  
\begin{align}
d\ln(P_t)=\mu_tdt+\sigma_tdw_t+\kappa_tdq_t
\end{align}

\begin{align}
RV_t=\sum^{1/\Delta t}_{j=1}r^2_{t-1+j\Delta t} \xrightarrow{p} \int_{t-1}^t\sigma_s^2ds\ + \sum_{s=1}^{q_t}\kappa_s^2= QV_t=IV_t+\sum_{s=1}^{q_t}\kappa_s^2 \ as\ \Delta t \to 0
\end{align}
In addition to the fact that volatility is a direct measure of the risk of portfolios, it is also important for pricing derivatives, such as options, since it is the only unknown parameter to estimate. The famous Black-Scholes model assumes a constant volatility over time which does not hold in practice. Implied volatility allows us to incorporate the jumps and continuous variation of volatility over time. In other words, implied volatility is an estimate of QV extracted from option prices.

VIX index estimates QV that is implied by option prices under a risk-neutral (18) process, using a finite number of strikes of S&P500 index [reference from CBOE]. It acknowledges the fact that assumptions of Black-Scholes model do not hold and the occurance of jumps. Since both VIX and RV are estimations of QV, it is reasonable to model VIX with a HAR-RV model which is usually used for modelling RV in the literature.

## 4.2 VIX descriptive statistics

We collect daily VIX index data from (ref: CBOE) starting from 27 Aug, 1996 to 31 July, 2020. And the descriptive statistics of $log(VIX_t)$ is shown as below:

```{r,message=F}
# data cleaning and import 
vix <- read.csv("^VIX.csv")
vix <- vix %>% dplyr::select("Date","Adj.Close")

vix$logvix <- log(vix$Adj.Close)

# previous day log vix values
prev <- matrix(0,nrow = nrow(vix)+1,ncol = 1)
for (i in 1:nrow(vix)){
  prev[1,1] <- c("NA")
  prev[i+1,1] <- vix$logvix[i]
}
prev <- as.data.frame(prev)
prev <- prev[-c(1:188,6189),1]

# need to be very careful about the time index -- dependent variable starting with t+1
MA5 <- matrix(0,nrow=nrow(vix)-5,ncol = 1)
for (i in 1:(nrow(vix)-5)){
  sum = 0
  for (j in 1:5){
    sum = sum + vix$logvix[i+j-1]
  }
  MA5[i,1] = (1/5)*sum
}

MA22 <- matrix(0,nrow = nrow(vix)-22,ncol = 1)
for (i in 1:(nrow(vix)-22)){
  sum = 0
  for (j in 1:22){
    sum = sum + vix$logvix[i+j-1]
  }
  MA22[i,1] = (1/22)*sum
}

#require(zoo)
#MA5 <- rollmean(vix$logvix,5,align = "center")
MA5 <- as.data.frame(MA5) # average of past week
MA5 <- MA5[-c(1:183),1] # including only 6000 obs

#MA22 <- rollmean(vix$logvix,22,align = "center")
MA22 <- as.data.frame(MA22) # average of past month
MA22 <- MA22[-c(1:166),1] # including only 6000 obs

logvix <- vix[-c(1:188),c(1,3)] # delete the adjusted vix col, only keep 6000 obs
logvix <- cbind(logvix,prev,MA5,MA22)
# export the dataset to csv. including only trading days and log returns

temp <- logvix[,2:5]
write_csv(temp,"vixlog.csv")

# the logvix, previous log vix and moving average values have the same length in the final file. 
```

```{r, results='hide'}
# descriptive statistics 
summary <- summary(vix$logvix) # ----> need to be tabulated 
# range(vix$logvix)
# some test results aginst normal dist h_null
# JB test
jb.norm.test(vix$logvix)
# Ljung box test
Box.test(vix$logvix,lag = 2,type = "Lj",fitdf = 1)

vixsk <- skewness(vix$logvix)
vixkt <- kurtosis(vix$logvix)
```

```{r}
sum <- data.frame(Series="Log(VIX)",Min = as.numeric(summary[1]), Median = as.numeric(summary[3]), Mean = as.numeric(summary[4]), Max = as.numeric(summary[6]),Skewness= vixsk,Kurtosis=vixkt,JB.Test=410.77,LB.Test=11765)
sum %>% 
  kable("latex",booktabs = T) %>% 
  kable_styling("bordered")%>% 
  add_header_above(c("Table ?: Descriptive statistics"=9))
```

```{r,message=FALSE,fig.width=10,fig.height=5}

# time series plot
logvix$Date <- as.Date(logvix$Date,format = "%Y-%m-%d")

p1 <- logvix %>% ggplot(aes(x = Date, y = logvix)) +
  geom_line()+theme_classic()+labs(x="Year",y="Daily log(VIX) values")+
  #ggtitle("Figure ?: Daily values for Log(VIX)")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  #scale_y_continuous(labels = scales::percent) + 
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=2,label="Global Financial Crisis",color="red",size=3)+
  annotate(geom = 'text',x=as.Date("2020-04-01"),y=2,label="COVID-19",color="red",size=3)+
  theme_clean()

# histogram plots
p2 <- logvix %>% ggplot()+geom_histogram(aes(x=logvix))+
  theme_clean()

# ACF plots to see the autocorrelation 
logvix <- logvix %>% 
  mutate(trading_day = row_number()) %>%
  as_tsibble(index = trading_day,regular = TRUE)

p3 <- logvix %>% ACF((logvix)^2) %>% autoplot()+
      xlab("number of lags")+ylab("ACF")+theme_clean()

figure <- ggarrange(p1,                                                 
          ggarrange(p2, p3, ncol = 2, labels = c("B", "C")), 
          nrow = 2, 
          labels = "A"                                        
          ) 
annotate_figure(figure, 
                top = text_grob("Descriptive figures",face = "bold"),
                fig.lab = "Figure 13",fig.lab.face = "bold"
                )
```

Table ? and Figure 13 shows that $log(VIX_t)$ is volatile and shows persistence over time. It is positively-skewed and the autocorrelation of its volatility is slowly decaying. From the descriptive statistics, we can see that the features of VIX is similar to what we have observed from S&P500. Similarly, there are unusual jumps happened during GFC and COVID-19. Thus we expect that focused score could outperform MLE and CRPS in predicting tails as before and significantly improve accuracy after GFC.

## 4.3 Model specification

Proposed by Corsi,(2003), HAR-RV model is a simple additive linear model which takes lagged squared returns as regressors. It does not belong to the class of long memory models, but it is able to produce the volatility persistence that is almost indistinguishable from what observed in financial markets through the simple AR-type structure.

We design the VIX analysis with different error term specification of HAR-RV:

1. $z_t \sim N(0,1)$

2. $z_t \sim Student-t(\mu,\sigma,\nu)$

3. $z_t \sim$ skewed Student-t distribution

4. $z_t \sim GARCH$ 
\begin{align}
log(VIX_{t+1})=\beta_0+\beta_1log(VIX_t)+\beta_2log(VIX_{t-5,t})+\beta_3log(VIX_{t-22,t})+\sigma_{t+1}z_{t+1}
\end{align}
\begin{align}
log(VIX_{t-5,t})=\frac{1}{5}[log(VIX_t)+...+log(VIX_{t-4})]
\end{align}
\begin{align}
log(VIX_{t-22,t})=\frac{1}{22}[log(VIX_t)+...+log(VIX_{t-21})]
\end{align}

By using the same methodology in section 2 with initial sample size $\tau = 1000$ and out-of-sample size $T-\tau+1=5000$, we produce the following average score table, Table ? and score trace plot, Figure ?.

### 4.3.1 Results of specification 1: Normal error

```{r}
vix_table <- read.csv("vix_table.csv")
vix_table <- rename(vix_table, "In-sample optimizers" = Row,"FSR10" = FSR10, "FSR20"=FSR20, "FSR80"=FSR80,"FSR90"=FSR90)
vix_table$`In-sample optimizers`<-c("LS","CRPS","FSR10","FSR20","FSR80","FSR90")

vix_table %>% 
  mutate(
    LS = cell_spec(format(LS,4),"latex",bold = ifelse(LS == max(LS),TRUE,FALSE)),
    CRPS = cell_spec(format(CRPS,4),"latex",bold = ifelse(CRPS==max(CRPS),TRUE,FALSE)),
    `FSR10` = cell_spec(format(`FSR10`,4),"latex",bold = ifelse(`FSR10`==max(`FSR10`),TRUE,FALSE)),
    `FSR20` = cell_spec(format(`FSR20`,4),"latex",bold = ifelse(`FSR20`==max(`FSR20`),TRUE,FALSE)),
    `FSR80` = cell_spec(format(`FSR80`,4),"latex",bold = ifelse(`FSR80`==max(`FSR80`),TRUE,FALSE)),
    `FSR90` = cell_spec(format(`FSR90`,4),"latex",bold = ifelse(`FSR90`==max(`FSR90`),TRUE,FALSE))
  ) %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F)%>%
  add_header_above(c("","Average out-of-sample scores"=6)) %>% 
  add_header_above(c("Table ?: Average scores for VIX"=7))
```

From Table ?, the results are not surprising focused score performs better in predicting upper tail as expected and does a relatively good job in predicting lower tail as well, even though CRPS ranks the best in terms of forecasting 20 percentile of the distribution.

The score trace plot, Figure ?, shows the prediction of optimizing four scoring rules evaluated by FSR10 (left panel) and FSR90 (right panel). It indicates how accurate these predictions are in 10 percentile and 90 percentile. The scoring rules used for optimization are {MLE, CRPS, FSR10(90), FSR20(80)}.

In the lower tail, it seems that there is not much difference among four predictions, but if zoom in, we can see that FSR is at the top of other scores, which means it can still provide the most accurate prediction. However, we only gain very little from using it. On the contrary, in the upper tail, FSR has a much higher effect on improving accuracy after GFC and this effect continues. 
One possible reason could be that considering the marginal distribution shown in figure 13.B of $log(VIX_t)$, both LS and CRPS can predict the lower tail better if there is higher probability mass at that end. Thus, the benefits of using FSR is not significant.

```{r,fig.width=10,fig.height=5}
# to compress all the calculated plotting average scores together in the same df 
trace_fsr10 <- read.csv("vixfsr10.csv")
trace_fsr20 <- read.csv("vixfsr20.csv")
trace_10ls <- read.csv("vixfsr10_ls.csv")
trace_10crps <- read.csv("vixfsr10_crps.csv")

fsr10 <- data.frame(trace_fsr10,trace_fsr20,trace_10ls,trace_10crps)

# add back dates to the fsr10 df, initial step = 1000

fsr10 <- fsr10 %>% mutate(Date = vix$Date[1001:6000])

# plot scores --- lower tail 
fsr10$Date <- as.Date(fsr10$Date,format = "%Y-%m-%d")

color1 <- c("FSR10"="red","FSR20"="black","MLE"="purple","CRPS"="green")

fsr10plot <- fsr10 %>% ggplot(aes(x=Date)) +
  geom_line(aes(y = FSR10,color = "FSR10"),size=0.25)+
  geom_line(aes(y = FSR20,color = "FSR20"),size=0.25)+
  geom_line(aes(y = MLE,color = "MLE"),size=0.25)+
  geom_line(aes(y = CRPS,color = "CRPS"),size=0.25)+
  labs(x="Year",y="Average Scores",color="legend")+
  #ggtitle("Figure 12: S&P500 Score traces-lower tail")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  scale_color_manual(values = color1)+
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=0.08,label="Global Financial Crisis",color="darkred")+
  theme_clean()

# plot scores --- upper tail 
trace_fsr90 <- read.csv("vixfsr90.csv")
trace_fsr80 <- read.csv("vixfsr80.csv")
trace_90ls <- read.csv("vixfsr90_ls.csv")
trace_90crps <- read.csv("vixfsr90_crps.csv")

fsr90 <- data.frame(trace_fsr90,trace_fsr80,trace_90ls,trace_90crps)

# add back dates to the fsr10 df, initial step = 1000

fsr90 <- fsr90 %>% mutate(Date = vix$Date[1001:6000])

# plot scores --- lower tail 
fsr90$Date <- as.Date(fsr90$Date,format = "%Y-%m-%d")

color2 <- c("FSR90"="red","FSR80"="black","MLE"="purple","CRPS"="green")

fsr90plot <- fsr90 %>% ggplot(aes(x=Date)) +
  geom_line(aes(y = FSR90,color = "FSR90"),size=0.25)+
  geom_line(aes(y = FSR80,color = "FSR80"),size=0.25)+
  geom_line(aes(y = MLE,color = "MLE"),size=0.25)+
  geom_line(aes(y = CRPS,color = "CRPS"),size=0.25)+
  labs(x="Year",y="",color="legend")+
  #ggtitle("Figure 12: S&P500 Score traces-upper tail")+
  scale_x_date(date_breaks = "years",date_labels = "%y")+
  scale_color_manual(values = color2)+
  annotate(geom = 'text',x=as.Date("2008-09-18"),y=-0.05,label="Global Financial Crisis",color="darkred")+
  theme_clean()

# put two plots side by side 
ggarrange(fsr10plot,fsr90plot,ncol=2,legend = "bottom",labels = c("Figure ?(a) Lower Tail","Figure ?(b) Upper Tail"))

```

### 4.3.2 Results of specification 2: Student-t

# 5. Conclusion

Caveats: 
1. The VaR is conditional distribution while the focused score is unconditional distribution due to the computation burden.

2. The practical meaning of the VaR is more often focused on around 1%, while at this very small tail, the focused score gets less and less accurate since the observations are less, harder to focus.

\pagebreak 
# References

DON'T forget alphabetical order!!

cite packages used in the draft
```{r}
citation("formattable")
```

\pagebreak
# Appendix 
## Appendix 1: Predictive Densities

The predictive density plots provide a straightforward way to visualize how optimal probabilistic forecast matches the shape of true DGP distribution through the simulation. Here we provide 4 different ‘time points’ to show how these scoring rules work, especially the focused score in figure 3 to 8. The blue line is the marginal distribution from the true DGP, and these figures show that predictive density produced from optimizing LS fits the area with high probability mass the best, which is also what LS is designed for. However, it performs poorly in terms of predicting tails when the degree of model misspecification is high in both GARCH and copula examples. This is when we benefit from using focused score, and we can see from figure 5 and 8 that predictions produced from optimizing focused score aim to match the shape of two tails of the true distribution. They outperform predictions from using Maximum likelihood estimation (MLE).

```{r}
knitr::include_graphics("figure 3.pdf")
```

```{r}
knitr::include_graphics("figure 4.pdf")
```

```{r}
knitr::include_graphics("figure 5.pdf")
```

```{r}
knitr::include_graphics("figure 6.pdf")
```

```{r}
knitr::include_graphics("figure 7.pdf")
```

```{r}
knitr::include_graphics("figure 8.pdf")
```