### 2021-05-23 portfolio.Rmd
The portfolio management of VIX - trading strategy 1 and 2.

---
title: "Portfolio"
author: "Yuru"
date: "16/05/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
link-citations: yes
bibliography: portfolio_ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Assumed predictive model

\begin{equation} \label{eq:HAR2}
log(VIX_{t+1})=\beta_0+\beta_1log(VIX_t)+\beta_2log(VIX_{t-5,t})+\beta_3log(VIX_{t-22,t})+\sigma_{t+1}z_{t+1}
\end{equation}
\begin{equation}
\epsilon_t = \sigma_t z_t
\end{equation}

\begin{equation}
z_t \sim N(0,1)
\end{equation}

\begin{equation} \label{eq:arch}
\sigma_{t+1}^2=\alpha_0+\alpha_1 \epsilon_t^2+\alpha_2 \sigma_t^2
\end{equation}

\begin{equation} \label{eq:ma5}
log(VIX_{t-5,t})=\frac{1}{5}[log(VIX_t)+...+log(VIX_{t-4})]
\end{equation}
\begin{equation} \label{eq:ma22}
log(VIX_{t-22,t})=\frac{1}{22}[log(VIX_t)+...+log(VIX_{t-21})]
\end{equation}

# Trading strategy 1 (Taylor,2019)

The trading strategy return is adjusted by the transaction cost and is defined as:
$$f(R_{t+1},\omega_t,\tau_t)=\frac{1+\omega_tR_{t+1}}{1+\tau_t},$$
where $R_{t+1}$ is the the return of the asset at time t+1, $\tau_t = \kappa|\omega_t-\omega_{t-1}|$ is the total transaction costs and $\kappa$ is the proportional transaction cost associated with the trading in the market.

The investor can take a long position, a short position or no position based on the next-day forecasted returns and the transaction costs incurred. The investor will take a long position if a long position generates the highest trading strategy return - $f(R_{t+1},\omega_t,\tau_t)$, compared with taking a short position or no position. Meanwhile, everytime the position changes, there are transaction costs generated. The trading strategy that takes transaction costs into account is referred as $\tau$-informed strategy. If the investor incorrectly assumes the transaction cost is zero, then it is referred as $\tau$-ignorant strategy.

The performances of these trading strategies are compared with three benchmark strategies.

1. Cash-only strategy: $f(R_{t+1},0,0)$

2. Long-only strategy: $f(R_{t+1},1,0)$

3. Short-only strategy: $f(R_{t+1},-1,0)$

The difference between the trading strategy returns and benchmarks are used to produce the Sharpe ratios. The average Sharpe ratios are listed in the tables.

```{r,message=FALSE}
library(dplyr)
library(tidyverse)
library(kableExtra)
```

```{r}
names <- matrix(c("Daily","Annulized"),2,1)
names <- as.data.frame(names)

# for CLS10 prediction 
cls10_inf <- read.csv("data/sharpe_inf10.csv")
cls10_inf_y <- cls10_inf*sqrt(252) # annualize sharpe ratio
cls10 <- cls10_inf %>% summarise_if(is.numeric,mean) %>% rbind(cls10_inf_y %>% summarise_if(is.numeric,mean))
cls10 <- cbind(names, cls10)
cls10 <- rename(cls10," " = V1)
cls10_ig <- read.csv("data/sharpe_fsr1.csv")
cls10_ig_y <- cls10_ig*sqrt(252)
cls10_ig <- cls10_ig %>% summarise_if(is.numeric,mean) %>% rbind(cls10_ig_y %>% summarise_if(is.numeric,mean))
cls10_ig <- cbind(names, cls10_ig)
cls10_ig <- rename(cls10_ig," "=V1)

# for cls90 prediction
cls90_inf <- read.csv("data/sharpe_inf90.csv")
cls90_inf_y <- cls90_inf*sqrt(252)
cls90 <- cls90_inf %>% summarise_if(is.numeric,mean) %>% rbind(cls90_inf_y %>% summarise_if(is.numeric,mean))
cls90 <- cbind(names, cls90)
cls90 <- rename(cls90," " = V1)

cls90_ig <- read.csv("data/sharpe_fsr4.csv")
cls90_ig_y <- cls90_ig*sqrt(252)
cls90_ig <- cls90_ig %>% summarise_if(is.numeric,mean) %>% rbind(cls90_ig_y %>% summarise_if(is.numeric,mean))
cls90_ig <- cbind(names, cls90_ig)
cls90_ig <- rename(cls90_ig," "=V1)

# for LS prediction 
ls_inf <- read.csv("data/sharpe_inf_LS.csv")
ls_inf_y <- ls_inf*sqrt(252)
ls <- ls_inf %>% summarise_if(is.numeric,mean) %>% rbind(ls_inf_y %>% summarise_if(is.numeric,mean))
ls <- cbind(names, ls)
ls <- rename(ls," "=V1)

ls_ig <- read.csv("data/taylor_LS.csv")
ls_ig_y <- ls_ig*sqrt(252)
ls_ig <- ls_ig %>% summarise_if(is.numeric,mean) %>% rbind(ls_ig_y %>% summarise_if(is.numeric,mean))
ls_ig <- cbind(names, ls_ig)
ls_ig <- rename(ls_ig," "=V1)

cls10_ig %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T1:Tau ignorant Sharpe ratio (CLS10)"=4))

cls10 %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T2:Tau informed Sharpe ratio (CLS10)"=4))

cls90_ig %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T3:Tau ignorant Sharpe ratio (CLS90)"=4))

cls90 %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T4:Tau informed Sharpe ratio (CLS90)"=4))

ls_ig %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T5:Tau ignorant Sharpe ratio (LS)"=4))

ls %>%
  kable("latex",escape = F, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T6:Tau informed Sharpe ratio (LS)"=4))
```

# Trading strategy 2

To evaluate the economic significance of the optimal probabilistic forecasts, we adopt the trading strategy based on interval forecasts in @kon2008 and calculate the corresponding Sharpe ratios. The rational of their strategy is that if the VIX price at time $t$ is close to the upper (lower) bound of the next day's predictive interval, then the price is likely to rise (decrease). It is then suggested that the investor takes a long (short) position. If the price does not show a sign of increasing (decreasing), then hold the current position, which means do nothing. That is,

if $log(VIX_t) < (>) \frac{U_{t+1|t}(\alpha)+L_{t+1|t}(\alpha)}{2}$, then take a long (short) position;

if $log(VIX_t) = \frac{U_{t+1|t}(\alpha)+L_{t+1|t}(\alpha)}{2}$, then hold the position.

$U_{t+1|t}$ and $L_{t+1|t}$ are the upper and lower bounds of the conditional predictive interval, which is constructed from optimal probabilistic forecasts.

Since CLS10 and CLS90 can predict the upper and lower tails of the conditional predictive distribution better, $U_{t+1|t}$ and $L_{t+1|t}$ are determined by CLS90 and CLS10 separately. That is, the trading strategy based on the CLS optimal forecasts is determined by the values of $log(VIX_t)$ and $\frac{U_{t+1|t}^{CLS90}+L_{t+1|t}^{CLS10}}{2}$.

We bring in the interval score (IS) as an optimizers to produce the optimal forecasts as in Section 3.1. By comparing the performance of IS and CLS, it might give us a better idea if forecasters can generate economic benefits from using CLS. The IS is defined as in @gneiting2007:

$$S_{IS}(P^{t-1}_{\boldsymbol{\theta}},y_t)=(u-l)+\frac{2}{\alpha}(l-y_t)I(y_t<l)+\frac{2}{\alpha}(y_t-u)I(y_t>u),$$
where $u$ and $l$ are the $1-\frac{\alpha}{2}$ and $\frac{\alpha}{2}$ quantiles of the $(1-\alpha)$x$100\%$ interval forecast.

We calculate the Sharpe ratio associated with the above trading strategies, which is used to measure the risk-adjusted return and the Sharpe ratio is defined as:

$$Sharpe \ ratio = \frac{R^p_t - rf_t}{\sigma^p_t}$$ 
$$R_t^p = log(VIX_{t+1}) - log(VIX_t),$$
where $R_t^p$ is the portfolio return at time $t$, $\sigma_t^p$ is the volatility of the portfolio at time $t$, and $rf_t$ is the risk-free rate at time $t$, which is approximated by the 10-year treasury constant maturity rate.

When the short position is taken at time $t$,
$$R_t^p = log(VIX_t)-log(VIX_{t+1})$$
The table below shows the daily and annualized average Sharpe ratios for the CLS, LS and IS optimal forecasts.

```{r}
cls_sharpe <- read.csv("data/sharpe_cls.csv")
cls_sharpe_y <- cls_sharpe*sqrt(252)
cls <- cls_sharpe %>% summarise_if(is.numeric,mean) %>% rbind(cls_sharpe_y %>% summarise_if(is.numeric,mean))
cls <- cbind(names, cls)
cls <- rename(cls," "=V1)

ls_sharpe <- read.csv("data/sharpe_LS.csv")
ls_sharpe_y <- ls_sharpe*sqrt(252)
ls_sharpe <- ls_sharpe %>% summarise_if(is.numeric,mean) %>% rbind(ls_sharpe_y %>% summarise_if(is.numeric,mean))
ls_sharpe <- cbind(names, ls_sharpe)
ls_sharpe <- rename(ls_sharpe," "=V1)

df <- cls

df %>% mutate(Sharpe_ls <- ls_sharpe) %>%
  kable("latex",escape = T, booktabs = T, linesep = "") %>%
  kable_styling("bordered",full_width = F) %>%
  add_header_above(c("T7: Sharpe ratio"=3))
```

# References